ActiveCA: Time Use Data from the General Social Survey of Canada to Study Active Travel

### 2025-05-29 paper/Manuscript-Data-Package-1Round-Review.tex
Editions for the first round of revision for the data paper.

\begin{abstract}
This paper describes \{ActiveCA\}, an open data product with Canadian
time use data. \{ActiveCA\} is an \texttt{R} data package that contains
analysis-ready data related to active travel spanning almost 40 years,
extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
(2010), and 29 (2015) of Canada's General Social Survey. Active travel
is characterized by mode, with walking being part of every cycle and
bicycling starting in 1992. The attributes of active trips are the types
of locations of origins and destinations, the duration of trips, and
episode weights for expanding the trips to population-wide estimates.
Based on the year of the survey, a variety of locations are coded. In
earlier cycles, these include home, work or school, and other's home,
whereas in later cycles these are augmented with locations such as
grocery stores, restaurants, outdoor destinations, and others. The
geographical resolution includes the province and whether the episode
was in an urban or rural setting.
\end{abstract}

@@

The objective of this paper is to introduce \{ActiveCA\}, an open data
product with time use data from Canada's General Social Surveys. Open
data products (ODPs) are the outcome of a process that transforms raw
data (open or not) into analysis-ready data, following a transparent
process in which all stages of development follow open principles
\citep{arribas-bel2021}. ODPs, while still open, differ from general
open data in their degree of ease of access, their heightened usability,
and potentially the value they add to the raw data.

\{ActiveCA\} provides analysis-ready data concerning active travel in
Canada spanning a period of almost 40 years. The source of these data is
Canada's program of General Social Surveys (GSSs). This program is
designed to provide cross-sectional data on topics of interest to
improve the well-being of Canadians. As part of this program, every five
to seven years the survey is done on the topic of time use. Concretely,
\{ActiveCA\} covers Cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
(2010), and 29 (2015) of the GSS. Time use data in these surveys is
coded using a very fine grain, from time spent in chores, leisure, and
sleeping, to time spent working or at school. These surveys have proved
valuable in investigations of mobility and quality of life
\citep{spinneyTransport2009}, the relationship between active travel and
transit use \citep{lachapelleLonger2016}, and travel behavior and time
poverty \citep{kimFacing2024}, to name but a few examples.

For \{ActiveCA\} and using GSS Public Use Microdata Files (PUMFs), we
extracted all data needed to characterize active travel in Canada,
namely, episodes where the activity was identified as moving between an
origin and a destination, either by walking or cycling. Although
Statistics Canada offers Public Use Microdata Files and documentation
for the GSS program \citep[see][]{statisticscanada2024}, accessing these
files, and preparing them for analysis is not a straightforward matter,
given their size and complexity. The process of extracting information
of interest from the source files is time-consuming, tedious, and
challenging and/or prone to error due to the expertise required to work
with these files. To create \{ActiveCA\} we collected, cleaned, and
processed the cycles of the GSS surveys concerning time use to make them
ready for analysis.

@@

\section{General Social Survey (GSS)
collection}\label{general-social-survey-gss-collection}

@@

The topics of the survey cycle every few years to cover topics that
include family, health, social identity, and every five to seven years
time use. The first Canadian time use survey done as part of the GSS
program was conducted in 1986, and the most recent was completed in
2015. These Time Use Surveys \citep{statisticscanada2022} collect data
on respondents' participation and time spent on a wide range of everyday
activities using a 24-hour retrospective diary, with information on the
location of these activities (e.g.~at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity. In addition, time-use surveys also cover
topics related to leisure time, work-life balance, health, commuting,
culture and sports, and many others.

@@

As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
labeled. Adittionaly, the format of the tables (comma-separated values)
does not allow for the specification of variable types (whether a
variable is continuous or discrete), which can lead to mistakes analysts
who have limited experience working with PUMFs.

@@

Figure \ref{fig:walking_documentation} displays the documentation for
the walking episodes from the 2015 GSS (\texttt{walking\_2015.rda}). The
documentation provides a brief description of the dataset, instructions
on how to load the package, as well as details about the file format and
descriptions of all variables (already labeled).

@@

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2015-documentation}

@@

\caption{Documentation for the dataset of walking episodes from 2015.}\label{fig:walking_documentation}
\end{figure}

The code below shows how to load the walking episodes from the 2015 GSS
(Cycle 29):

\begin{verbatim}
data(walking_2015)
\end{verbatim}

@@

For instance, to calculate the percentage of respondents from the 2015
GSS survey with active travel episodes, it is necessary to account for
the person's weight. In 2015, the weight variable is represented by
\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
percentage of people with active travel episodes by age group. It uses
the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.

The process begins by creating a dataset that sums the population by age
group. Then, it joins the 2015 episodes with the 2015 Main File. Note
that in both operations, the code sums the person weight variable
(\texttt{WGHT\_PER}) to obtain the correct population values. After
this, a new dataset, \texttt{Active\_percentage}, is created by merging
both previous datasets. The percentage of active travel episodes by age
group is calculated by dividing the total population by the population
with active trip episodes, then multiplying by 100 and rounding with 2
decimal places.

@@

# Calculate the total population for each age group
Total_population <- gss_main_2015 |> 
  group_by(AGEGR10) |> 
  summarise(Total_population  = sum(WGHT_PER))

# Calculate the active population for each age group
Active_population <- gss_episodes |> 
      filter(YEAR == 2015) |> 
  left_join(gss_main_2015, by = c("PUMFID"="PUMFID")) |> 
  group_by(AGEGR10) |> 
  summarise(Active_population  = sum(WGHT_PER))

@@

The result shows that the age group with the highest share of active
trips is those between 15 and 20 years old, with almost 37\%, followed
by those between 25 and 34 years old with around 33\%. There is a
significant drop in percentage for the following groups, with the
percentage falling to between 15\% and 18\% for the remaining age
groups.

\begin{longtable}[]{@{}lrrr@{}}
\toprule\noalign{}
AGEGR10 & Total\_population & Active\_population & Percentage \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
15 to 24 years & 4511131 & 1641468.3 & 36.39 \\
25 to 34 years & 4956386 & 1644032.8 & 33.17 \\
35 to 44 years & 4734506 & 847769.9 & 17.91 \\
45 to 54 years & 5136125 & 790667.3 & 15.39 \\
55 to 64 years & 4831306 & 767064.0 & 15.88 \\
65 to 74 years & 3283969 & 559024.6 & 17.02 \\
75 years and over & 2312976 & 384358.6 & 16.62 \\
\end{longtable}

@@

Table \ref{tab:processed-obs} displays the total number of records
processed for Main and Episode Files. For the Main Files, a total of
89,331 records were processed, referring to all records from the Time
Use Surveys from 1986 to 2015, that together represents more of
149,389,839 respondents (29,766,399 for 2015, 28,075,610 for 2010,
26,095,819 for 2005, 24,260,137 for 1998, 21,294,313 for 1992, and for
2005, 24,260,137 for 1998, and 21,294,313 for 1986). It also presents
the total cases of active trips episodes identified. In total 21,748
records with register of active travel activity. Together, these records
account for 38,275,077 episodes.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{lrrr}
\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{2}{c}{Observations} \\
\cmidrule(l{3pt}r{3pt}){3-4}
Survey & Year & Count & Weighted\\
\midrule
 & 2015 & 17390 & 29766399\\
\nopagebreak
 & 2010 & 15390 & 28075610\\
\nopagebreak
 & 2005 & 19597 & 26095819\\
\nopagebreak
 & 1998 & 10749 & 24260137\\
\nopagebreak
 & 1992 & 9815 & 21294313\\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash Main} & 1986 & 16390 & 19897562\\
\cmidrule{1-4}\pagebreak[0]
 & 2015 & 3496 & 6634387\\
\nopagebreak
 & 2010 & 4615 & 8516753\\
\nopagebreak
 & 2005 & 5866 & 7583838\\
\nopagebreak
 & 1998 & 1789 & 3606987\\
\nopagebreak
 & 1992 & 1635 & 3691918\\
\nopagebreak
\multirow[t]{-6}{*}{\raggedright\arraybackslash Episode} & 1986 & 4347 & 8241196\\
\bottomrule
\end{longtable}
\endgroup{}

@@

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `WGHT\_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the Main File of the 2015 GSS.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
\addlinespace
10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

Considering GSS Cycles analyzed, we identified 21,748 episodes that
recorded active travel episodes, with trip duration ranging from 0 to
900 minutes, to twelve different destinations. \{ActiveCA\} includes all
these episodes ready for analysis. Table \ref{tab:table-01} presents
descriptive statistics on walking and cycling trips between 1986 and
2015, with measures of trip duration in minutes. The 1986 survey did not
include bicycle trips.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{>{}llcccccc}
\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{6}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-8}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015\\
\midrule
 & Maximum & 660 & 300 & 255 & 515 & 480 & 900\\
\nopagebreak
 & Mean & 21 & 21 & 12 & 12 & 13 & 18\\
\nopagebreak
 & Median & 15 & 10 & 5 & 10 & 10 & 10\\
\nopagebreak
 & Minimum & 1 & 1 & 1 & 0 & 0 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27\\
\cmidrule{1-8}\pagebreak[0]
 & Maximum &  & 240 & 90 & 180 & 153 & 120\\
\nopagebreak
 & Mean &  & 28 & 24 & 20 & 19 & 25\\
\nopagebreak
 & Median &  & 15 & 15 & 15 & 10 & 20\\
\nopagebreak
 & Minimum &  & 5 & 2 & 1 & 1 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:table-01} shows that, in general, the median values for
walking trips is 10 minutes, except for 1998 when the median was 5
minutes. In the case of cycling trips, the duration fluctuated over the
years, ranging from 10 to 20 minutes. The table also highlights very
high maximum values, particularly for walking trips, with recorded
episodes exceeding 4 hours in all cases.

Table \ref{tab:table-02} and \ref{tab:table-03} provide descriptive
statistics for the two modes of transportation, split by destination
categories, from 1986 to 1998 and from 2005 to 2015, respectively. In
Table \ref{tab:table-02}, one can observed that in 1986 and 1992,
walking trips destined for \texttt{home} had the highest medians.
However, by 1998, the highest medians shifted to trips to
\texttt{work\ or\ school}, a transition that also occurred for cycling
trips between 1992 and 1998. Table \ref{tab:table-03} indicates that the
median duration for walking trips to \texttt{home} and
\texttt{work\ or\ school} remained at 10 minutes.

\begingroup\fontsize{6}{8}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item 'Min' denotes the minimum time to reach the destination; 'Max' denotes the maximum time to reach the destination; '(\%)' indicates a percentage of the total time to reach the destination; 'Med' refers to the median time to reach the destination
\end{TableNotes}
\begin{longtable}[t]{ccccc>{}c|ccc>{}c|cccc}
\caption{\label{tab:table-02}\label{tab:table-02}Comparison of travel statistics by mode and destination: 1986, 1992, 1998}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{4}{c}{1986} & \multicolumn{4}{c}{1992} & \multicolumn{4}{c}{1998} \\
\cmidrule(l{3pt}r{3pt}){3-6} \cmidrule(l{3pt}r{3pt}){7-10} \cmidrule(l{3pt}r{3pt}){11-14}
\multicolumn{1}{c}{\textbf{Destination}} & \multicolumn{1}{c}{\textbf{Mode}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}} & \multicolumn{1}{c}{\textbf{Min}} & \multicolumn{1}{c}{\textbf{Med}} & \multicolumn{1}{c}{\textbf{Max}} & \multicolumn{1}{c}{\textbf{(\%)}}\\
\midrule
 & Home &  &  &  &  & 5 & 20 & 240 & 55.7 & 2 & 15 & 90 & 51.6\\
\nopagebreak
 & Other's home &  &  &  &  & 5 & 10 & 145 & 19.7 & 2 & 10 & 80 & 15.7\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Cycling} & Work or school &  &  &  &  & 5 & 15 & 45 & 24.7 & 5 & 25 & 75 & 32.8\\
\cmidrule{1-14}\pagebreak[0]
 & Home & 1 & 15 & 330 & 46.3 & 1 & 15 & 300 & 60.5 & 1 & 5 & 255 & 51.9\\
\nopagebreak
 & Other's home & 1 & 10 & 660 & 42.6 & 1 & 5 & 135 & 18.7 & 1 & 5 & 120 & 27.8\\
\nopagebreak
\multirow[t]{-3}{*}{\centering\arraybackslash Walking} & Work or school & 1 & 10 & 450 & 11.1 & 2 & 10 & 60 & 20.9 & 1 & 10 & 75 & 20.3\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

\begingroup\fontsize{6}{8}\selectfont

>>>>

\begin{abstract}
This paper describes \{ActiveCA\}, an open data product with Canadian
time use data. \{ActiveCA\} is an \texttt{R} data package that contains
analysis-ready data related to active travel spanning almost 40 years,
extracted from cycles 2 (1986), 7 (1992), 12 (1998), 19 (2005), 24
(2010), 29 (2015), and 34 (2022) of the Time Use Survey (TUS) cicles
from the General Social Survey (GSS). Active travel is characterized by
mode, with walking being part of every cycle and bicycling starting in
1992. The attributes of active trips are the types of locations of
origins and destinations, the duration of trips, and episode weights for
expanding the trips to population-wide estimates. Based on the year of
the survey, a variety of locations are coded. In earlier cycles, these
include home, work or school, and other's home, whereas in later cycles
these are augmented with locations such as grocery stores, restaurants,
outdoor destinations, and others. The geographical resolution includes
the province and whether the episode was in an urban or rural setting.
\end{abstract}

@@

The objective of this paper is to introduce \{ActiveCA\}, an open data
product with data from all Time Use Survey (TUS) cycles of the General
Social Surveys (GSSs). Open data products (ODPs) are the outcome of a
process that transforms raw data (open or not) into analysis-ready data,
following a transparent process in which all stages of development
follow open principles \citep{arribas-bel2021}. ODPs, while still open,
differ from general open data in their degree of ease of access, their
heightened usability, and potentially the value they add to the raw
data.

\{ActiveCA\} provides analysis-ready data concerning active travel in
Canada spanning a period of almost 40 years. The source of these data is
TUS from the GSS. This program is designed to provide cross-sectional
data on topics of interest to improve the well-being of Canadians. As
part of this program, every five to seven years the survey is done on
the topic of time use. Concretely, \{ActiveCA\} covers Cycles 2 (1986),
7 (1992), 12 (1998), 19 (2005), 24 (2010), 29 (2015), and 34 (2022) of
the GSS. Time use data in these surveys is coded using a very fine
grain, from time spent in chores, leisure, and sleeping, to time spent
working or at school. These surveys have proved valuable in
investigations of mobility and quality of life
\citep{spinneyTransport2009}, the relationship between active travel and
transit use \citep{lachapelleLonger2016}, and travel behavior and time
poverty \citep{kimFacing2024}, to name but a few examples.

For \{ActiveCA\} and using TUS - GSS Public Use Microdata Files (PUMFs),
we extracted all data needed to characterize active travel in Canada,
namely, episodes where the activity was identified as moving between an
origin and a destination, either by walking or cycling. Although
Statistics Canada offers Public Use Microdata Files and documentation
for the GSS program \citep[see][]{statisticscanada2024}, accessing these
files, and preparing them for analysis is not a straightforward matter,
given their size and complexity. The process of extracting information
of interest from the source files is time-consuming, tedious, and
challenging and/or prone to error due to the expertise required to work
with these files. To create \{ActiveCA\} we collected, cleaned, and
processed the cycles of the GSS surveys concerning time use to make them
ready for analysis.

@@

\section{The Time Use Survey (TUS)
collection}\label{the-time-use-survey-tus-collection}

@@

The topics of the survey cycle every few years to cover topics that
include family, health, social identity, and every five to seven years
time use (TUS). The first Canadian TUS done as part of the GSS program
was conducted in 1986, and the most recent was completed in 2022. These
Time Use Surveys \citep{statisticscanada2022} collect data on
respondents' participation and time spent on a wide range of everyday
activities using a 24-hour retrospective diary, with information on the
location of these activities (e.g.~at home, at work, etc.) and, for
non-personal activities, the people who were present with the respondent
at the time of the activity. In addition, time-use surveys also cover
topics related to leisure time, work-life balance, health, commuting,
culture and sports, and many others.

TUS allows researchers to identify the origin and destination of trips,
travel times and modes of transport used, providing a valuable dataset
for analyzing active travel behavior. It also provides the empirical
basis for tools used in transportation analysis, such as the development
of impedance functions for accessibility analysis, a measure of the ease
with which people can reach destinations and opportunities
\citep{hansen1959, paez2012}. The Canadian TUS is unique at the national
level in collecting detailed information on travel behaviour. Its
consistent application across survey cycles enables the identification
of long-term trends, with some questions present in the questionnaires
since the first application of the survey.

Most respondents to the 2022 TUS completed it online, reflecting
Statistics Canada's effort to adapt to technological changes and growing
time demands by offering greater flexibility and convenience
\citeyearpar{statisticscanada2022}. While such methodological changes
may affect data comparability over time, it is not possible to determine
whether observed differences result from actual population changes or
shifts in data collection methods. Despite rigorous efforts to ensure
data quality, the use of electronic questionnaires may have influenced
estimates. Statistics Canada assessed the impact of collection mode on a
limited set of key questions, constrained by sample size. Importantly,
none of the variables used in this research are flagged in the 2022 PUMF
User Guide as unsuitable for trend analysis.

Until 2022, Statistics Canada employed a telephone-based sampling frame,
which was replaced by a dwelling-based frame in the most recent cycle.
Each survey cycle spans a 12-month period, typically from July to the
following July. The target population includes all Canadians aged 15 and
over, excluding residents of the Yukon, Northwest Territories, and
Nunavut, full-time institutional residents, and individuals living on
Indigenous reserves.

The survey encompasses both rural and urban areas, including
metropolitan and non-metropolitan regions, to ensure a diverse and
representative sample. For sampling, the ten provinces were divided into
geographic strata. Several Census Metropolitan Areas (CMAs) - such as
St.~John's, Halifax, Saint John, Montreal, Quebec City, Toronto, Ottawa,
Hamilton, Winnipeg, Regina, Saskatoon, Calgary, Edmonton, and Vancouver
- were treated as separate strata. Additional strata grouped other CMAs
within Quebec, Ontario, and British Columbia, as well as non-CMA areas
within each province.

@@

As shown in Table \ref{tab:main-2015-unprocessed}, the variables are not
labeled. Additionally, the format of the tables (comma-separated values)
does not allow for the specification of variable types (whether a
variable is continuous or discrete), which can lead to mistakes analysts
who have limited experience working with PUMFs.

@@

Figure \ref{fig:walking_documentation} displays the documentation for
the walking episodes from the 2022 GSS (\texttt{walking\_2022.rda}). The
documentation provides a brief description of the dataset, instructions
on how to load the package, as well as details about the file format and
descriptions of all variables (already labeled).

@@

{\centering \includegraphics[width=1\linewidth]{Manuscript-figures/walking-2022-documentation}

@@

\caption{Documentation for the dataset of walking episodes from 2022.}\label{fig:walking_documentation}
\end{figure}

The code below shows how to load the walking episodes from the 2022 GSS
(Cycle 29):

\begin{verbatim}
data(walking_2022)
\end{verbatim}

@@

For instance, to calculate the percentage of respondents from the 2022
GSS survey with active travel episodes, it is necessary to account for
the person's weight. In 2022, the weight variable is represented by
\texttt{WGHT\_PER.} The code below demonstrates how to obtain the
percentage of people with active travel episodes by age group. It uses
the \texttt{dplyr} package \citep{dplyr2023} to manipulate the data.

The process begins by creating a dataset that sums the population by age
group. Then, it joins the 2022 episodes with the 2022 Main File. Note
that in both operations, the code sums the person weight variable
(\texttt{WGHT\_PER}) to obtain the correct population values. After
this, a new dataset, \texttt{Active\_percentage}, is created by merging
both previous datasets. The percentage of active travel episodes by age
group is calculated by dividing the total population by the population
with active trip episodes, then multiplying by 100 and rounding with 2
decimal places.

@@

# Calculate the total population for each age group
Total_population <- gss_main_2022 |> 
  group_by(AGEGR10) |> 
  summarise(Total_population  = sum(WGHT_PER))

# Calculate the active population for each age group
Active_population <- gss_episodes |> 
      filter(YEAR == 2022) |> 
  left_join(gss_main_2022, by = c("PUMFID"="PUMFID")) |> 
  group_by(AGEGR10) |> 
  summarise(Active_population  = sum(WGHT_PER))

@@

The result shows that the age group with the highest share of active
trips is those between 15 and 20 years old, with almost 44\%, followed
by those between 25 and 34 years old with around 19\%. There is a
significant drop in percentage for the following groups, with the
percentage falling to between 12\% and 15\% for the remaining age
groups.

\begin{longtable}[]{@{}lrrr@{}}
\toprule\noalign{}
AGEGR10 & Total\_population & Active\_population & Percentage \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
15 to 24 years & 4537139 & 1984658.8 & 43.74 \\
25 to 34 years & 5364748 & 1048653.8 & 19.55 \\
35 to 44 years & 5227988 & 669688.8 & 12.81 \\
45 to 54 years & 4720990 & 605162.1 & 12.82 \\
55 to 64 years & 5143935 & 690336.5 & 13.42 \\
65 to 74 years & 4160941 & 619732.8 & 14.89 \\
75 years and over & 2981061 & 422799.5 & 14.18 \\
\end{longtable}

@@

Table \ref{tab:processed-obs} displays the total number of records
processed for Main and Episode Files. For the Main Files, a total of
101,667 records were processed, referring to all records from the Time
Use Surveys from 1986 to 2022, that together represents more of
181,526,641 respondents (32,136,802 for 2022, 29,766,399 for 2015,
28,075,610 for 2010, 26,095,819 for 2005, 24,260,137 for 1998,
21,294,313 for 1992, and for 2005, 24,260,137 for 1998, and 21,294,313
for 1986). It also presents the total cases of active trips episodes
identified. In total 23,513 records with register of active travel
activity. Together, these records account for 44,316,110 episodes.

\begin{longtable}[t]{rr>{}r|rr}
\caption{\label{tab:table_df_processed}\label{tab:processed-obs}Total number and weighted sum of records processed.}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Main} & \multicolumn{2}{c}{Episode} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Year & Count & Weighted & Count\_ep & Weighted\_ep\\
\midrule
2022 & 12336 & 32136802 & 1765 & 6041032\\
2015 & 17390 & 29766399 & 3496 & 6634387\\
2010 & 15390 & 28075610 & 4615 & 8516753\\
2005 & 19597 & 26095819 & 5866 & 7583838\\
1998 & 10749 & 24260137 & 1789 & 3606987\\
\addlinespace
1992 & 9815 & 21294313 & 1635 & 3691918\\
1986 & 16390 & 19897562 & 4347 & 8241196\\
\bottomrule
\end{longtable}
\endgroup{}

@@

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item Legend: `PUMFID`: record identification. `WGHT\_PER`:  person weight. `SURVMNTH`: survey month of data collection. `AGEGR10`: age group of the respondent. `SEX`: sex of the respondent. `MARSTAT`: marital status of the respondent.
\end{TableNotes}
\begin{longtable}[t]{cccccc}
\caption{\label{tab:gss-processed-file-2015}\label{tab:main-2015-processed}Visualization of the first ten lines and first six columns of the 2015 TUS Main File.}\\
\toprule
PUMFID & WGHT\_PER & SURVMNTH & AGEGR10 & SEX & MARSTAT\\
\midrule
10000 & 616.6740 & July & 55 to 64 years & Male & Divorced\\
10001 & 8516.6140 & July & 55 to 64 years & Male & Married\\
10002 & 371.7520 & January & 45 to 54 years & Female & Married\\
10003 & 1019.3135 & March & 65 to 74 years & Female & Divorced\\
10004 & 1916.0708 & September & 25 to 34 years & Male & Single, never married\\
\addlinespace
10005 & 1952.2015 & April & 15 to 24 years & Male & Single, never married\\
10006 & 5761.5528 & August & 15 to 24 years & Male & Single, never married\\
10007 & 466.0426 & June & 55 to 64 years & Female & Widowed\\
10008 & 2479.2991 & February & 25 to 34 years & Female & Married\\
10009 & 1436.1641 & August & 65 to 74 years & Male & Widowed\\
\bottomrule
\insertTableNotes
\end{longtable}
\end{ThreePartTable}
\endgroup{}

@@

Considering all TUS analyzed, we identified 23,513 episodes that
recorded active travel episodes, with trip duration ranging from 0 to
900 minutes, to twelve different destinations. \{ActiveCA\} includes all
these episodes ready for analysis. Table \ref{tab:table-01} presents
descriptive statistics on walking and cycling trips between 1986 and
2022, with measures of trip duration in minutes. The 1986 survey did not
include bicycle trips.

\begin{longtable}[t]{>{}llccccccc}
\caption{\label{tab:table-01}\label{tab:table-01}Descriptive statistics for episodes with active transport records}\\
\toprule
\multicolumn{2}{c}{ } & \multicolumn{7}{c}{Year} \\
\cmidrule(l{3pt}r{3pt}){3-9}
Mode & Statistic & 1986 & 1992 & 1998 & 2005 & 2010 & 2015 & 2022\\
\midrule
 & Maximum & 660 & 300 & 255 & 515 & 480 & 900 & 480\\
\nopagebreak
 & Mean & 21 & 21 & 12 & 12 & 13 & 18 & 19\\
\nopagebreak
 & Median & 15 & 10 & 5 & 10 & 10 & 10 & 15\\
\nopagebreak
 & Minimum & 1 & 1 & 1 & 0 & 0 & 5 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Walking}} & Standard deviation & 31 & 25 & 17 & 16 & 17 & 27 & 24\\
\cmidrule{1-9}\pagebreak[0]
 & Maximum &  & 240 & 90 & 180 & 153 & 120 & 150\\
\nopagebreak
 & Mean &  & 28 & 24 & 20 & 19 & 25 & 40\\
\nopagebreak
 & Median &  & 15 & 15 & 15 & 10 & 20 & 30\\
\nopagebreak
 & Minimum &  & 5 & 2 & 1 & 1 & 5 & 5\\
\nopagebreak
\multirow[t]{-5}{*}{\raggedright\arraybackslash \textbf{Cycling}} & Standard deviation &  & 36 & 18 & 18 & 23 & 20 & 20\\
\bottomrule
\end{longtable}
\endgroup{}

Table \ref{tab:table-01} shows that, until 2022 the median values for
walking trips were 10 minutes, increasing to 15 minutes in the last
survey. In the case of cycling trips, the duration fluctuated over the
years, ranging from 10 to 30 minutes. The table also highlights very
high maximum values, particularly for walking trips, with recorded
episodes exceeding 4 hours in all cases.

\{ActiveCA\} also enables visual analysis of active travel in Canada
using traditional exploratory data analysis techniques. Figure
\ref{fig:figure-01} shows walking trips from 2022 through heat maps.
This graph uses color gradients to represent the percentage of trips
between various origins and destinations, with darker colors indicating
higher percentages and lighter colors representing less frequent routes.
For conciseness, we omitted the heat maps for the other years analyzed.

In 2022, \texttt{home} served as a central hub for most trips, with
fewer than 10\% of journeys not involving it as either a starting point
or destination. The most common trip types were from \texttt{home} to
\texttt{work\ or\ school} (17\%) and the reverse, from
\texttt{work\ or\ school} to \texttt{home} (13\%). Notably, 7\% of trips
began and ended at \texttt{home}, often reflecting leisure activities
such as short walks or dog walking. \texttt{Grocery\ stores} were also a
key destination, comprising 10\% of trips departing from \texttt{home.}
